Even-point Multi-loop Unitarity and its Applications: Exponentiation, Anomalies and Evanescence

### 2023-07-25 eventPointTwoLoop.tex
reword end of sec. 5

and the second is the series that counts {symmetric invariants}, $s_{ij}^x(s_{ik}s_{jk})^y$, which we denote as $\mathcal{H}^{(ij)(kl)}$,

@@

where we have defined the bracket $[\mathcal{H}]$ such that it maps Hilbert series with integer coefficients at successive orders in $\alpha$ into a sequence of numbers. Both of these Hilbert series are simple rational functions of $\alpha$, which we state below:

@@

With this, we can determine the number of evanescent operators that can contribute at each successive order in $\alpha'$ for four-photon matrix elements. The number of evanescent operators at each mass-dimension is just the difference between the dimension of the general dimension Hilbert series, and that of the $D=4$ Hilbert series, $\mathcal{H}^{\text{ev.}}=\mathcal{H}^{\text{gen.}D} - \mathcal{H}^{D=4}$. Thus we obtain the following Hilbert series for the number of evanescent four-photon operators at $\mathcal{O}(\alpha'^{n+2})$:

@@

In this manuscript, we have carried out a detailed analysis of even-point effective field theories at NNLO in the perturbative expansion. In \sect{sec:Review}, we provided a review of the generalized unitarity and integration methods employed throughout the text. Then in \sect{sec:EMU} we introduced and developed the on-shell constructive method of Even-point Multi-loop Unitarity (EMU), and computed the tensor reduction for triangle and bubble integrals in $D$-dimensions of arbitrary rank along with a spanning set of four-photon operators needed to capture higher-loop effects. Due to the simplicity of even-point multi-loop amplitudes of NLSM and DBIVA, these methods allowed us to compute fully integrated two-loop amplitudes for NLSM, pure Born-Infeld, and $\mathcal{N}=4$ DBIVA theory in \sect{sec:Loops}. Finally, in \sect{sec:Actions} we studied the quantum effective actions that capture the aforementioned loop effects. In doing so, we have identified a variety of rich physical structures that we summarize below: 
\paragraph{\textbf{Exponentiation}} In \eqn{eq:NLSM2bub} and \eqn{eq:NLSMostrich} we computed in general dimension, $D$, the two contributions to NLSM two-loop amplitudes. Plugging in explicit color structures, we found that the leading divergence of NLSM amplitudes on a $\mathbb{CP}^1$ target space exponentiate in $D=2-2\epsilon$, akin to the IR exponentiation of gravity amplitudes \cite{Weinberg:1965nx,Naculich:2008ew,White:2011yy,DiVecchia:2019myk,DiVecchia:2019kta,Heissenberg:2021tzo}. In addition, evaluating these diagrams in the planar limit, where $N_c \rightarrow \infty$, we found that the double-bubble integral iterates the leading divergence and subleading logarithms in the full one-loop amplitude. Thus, identifying corrections to the NLSM Lagrangian that absorb the ostrich diagram would imbue the scale-dependent logarithms with exponential structure at loop-level. This non-trivial property is found in theories that are conjectured to be integrable \cite{Shankar:1977cm,Zamolodchikov:1977nu,Zamolodchikov:1978xm, Komatsu:2019hgc}, like $\mathcal{N}=4$ super-Yang-Mills \cite{Anastasiou:2003kj,Bern:2005iz} in $D=4-2\epsilon$. In future work, we hope to study whether this iterative structure can be further applied beyond the $\mathbb{CP}^1$ model.

>>>>

and the second sequence counts {symmetric invariants}, $s_{ij}^x(s_{ik}s_{jk})^y$, which we denote as $\mathcal{H}^{(ij)(kl)}$,

@@

Above, we have defined the bracket $[\mathcal{H}]$ such that it maps Hilbert series with integer coefficients into a sequence of numbers at successive orders in $\alpha$. Both of these Hilbert series are simple rational functions of $\alpha$, which we state below:

@@

With this we can determine the number of four-photon evanescent operators that contribute at each successive order in $\alpha'$. The number of evanescent operators is determined by the difference in operator dimension between the general-$D$ Hilbert series and the $D=4$ Hilbert series, $\mathcal{H}^{\text{ev.}}=\mathcal{H}^{\text{gen.}D} - \mathcal{H}^{D=4}$. Thus we obtain the following Hilbert series for the number of evanescent four-photon operators at $\mathcal{O}(\alpha'^{n+2})$:

@@

In this manuscript, we have carried out a detailed analysis of even-point effective field theories through two-loop in the perturbative expansion. In \sect{sec:Review}, we provided a review of the generalized unitarity and integration methods employed throughout the text. Then in \sect{sec:EMU} we introduced and developed the on-shell constructive method of Even-point Multi-loop Unitarity (EMU), and computed the tensor reduction for triangle and bubble integrals in $D$-dimensions of arbitrary rank along with a spanning set of four-photon operators needed to capture higher-loop effects. Due to the simplicity of even-point multi-loop amplitudes of NLSM and DBIVA, these methods allowed us to compute fully integrated two-loop amplitudes for NLSM, pure Born-Infeld, and $\mathcal{N}=4$ DBIVA theory in \sect{sec:Loops}. Finally, in \sect{sec:Actions} we studied the quantum effective actions that capture the aforementioned loop effects. In doing so, we have identified a variety of rich physical structures that we summarize below: 
\paragraph{\textbf{Exponentiation}} In \eqn{eq:NLSM2bub} and \eqn{eq:NLSMostrich} we computed in general dimension, $D$, the two contributions to NLSM two-loop amplitudes. Plugging in explicit color structures, we found that the leading divergence of NLSM amplitudes on a $\mathbb{CP}^1$ target space exponentiate in $D=2-2\epsilon$, akin to the IR exponentiation of gravity amplitudes \cite{Weinberg:1965nx,Naculich:2008ew,White:2011yy,DiVecchia:2019myk,DiVecchia:2019kta,Heissenberg:2021tzo}. In addition, evaluating these diagrams in the planar limit, where $N_c \rightarrow \infty$, we found that the double-bubble integral iterates the leading divergence and subleading logarithms in the full one-loop amplitude. Thus, identifying corrections to the NLSM Lagrangian that absorb the ostrich diagram would imbue the scale-dependent logarithms with exponential structure at loop-level. This non-trivial property is found in theories that are conjectured to be integrable \cite{Shankar:1977cm,Zamolodchikov:1977nu, Komatsu:2019hgc}, like $\mathcal{N}=4$ super-Yang-Mills \cite{Anastasiou:2003kj,Bern:2005iz} in $D=4-2\epsilon$. In future work, we hope to study whether this iterative structure can be further applied beyond the $\mathbb{CP}^1$ model.
